---
title: "PM2.5 Exposure and Dementia Risk"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
output_file: index.html
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(plotly)
library(DT)
library(crosstalk)
library(crosstalk)
```

```{r data}
EHBS <- read.csv("C:/Users/keita/OneDrive - Emory University/Thesis - Anke Huels/Dataset/ISEEMarch19data2.csv")

```

### Age Distribution of Study Participants

This widget looks at the calculate age of the participants in the EHBS cohort at each visit. This will allow us to create a scale that is adjustable and see how many participants also lie within that age of age ranges. It is normally distributed, and will allow us to explore specific age range afefcts.

```{r}
# Shared data setup for filtering
shared_data <- SharedData$new(EHBS, ~pt_calculated_age)

# Age range slider
age_slider <- filter_slider(
  "age_range", "Select Age Range:",
  shared_data, ~pt_calculated_age,
  step = 1, width = "100%"
)

# Descriptive stats function
calculate_stats <- function(data) {
  age <- data$pt_calculated_age
  stats <- data.frame(
    Statistic = c("Mean", "Median", "Standard Deviation", "Minimum", "Maximum", "Number of Participants"),
    Value = c(
      mean(age, na.rm = TRUE),
      median(age, na.rm = TRUE),
      sd(age, na.rm = TRUE),
      min(age, na.rm = TRUE),
      max(age, na.rm = TRUE),
      length(age)
    )
  )
  return(stats)
}

# Interactive histogram
age_plot <- plot_ly(shared_data, x = ~pt_calculated_age, type = "histogram",
                    marker = list(color = 'rgba(50, 171, 96, 0.6)')) %>%
  layout(
    title = "Age Distribution of Participants",
    xaxis = list(title = "Age (years)"),
    yaxis = list(title = "Number of Participants")
  )

# Summary stats
stats_table <- calculate_stats(shared_data$data(withSelection = TRUE))

# Interactive table
stats_dt <- datatable(stats_table, options = list(pageLength = 6, dom = 't'), rownames = FALSE) %>%
  formatRound(columns = "Value", digits = 2)

# Layout
bscols(
  age_slider,
  bscols(age_plot, stats_dt, widths = c(6, 6))
)
```

### Forest Plot: Cognitive Tests and PM2.5 Exposure
This widget is looking at 5 different cognitive assessments and multi-year air pollution averages, which will allow us to explore the PM2.5 and Rline2.5 air pollution concentrations in the Emory Healthy Brain Study cohorts. We can see that most assocations are null, showing that in this cohort, there is no signfiicant assocation between air pollution exposure and cognitive function.

```{r}
# Load and format all test data
tests <- c("Delay Recall Test", "Recognition Test", "Number Span Test", 
           "Multilingual Naming Test", "Animal Fluency Test")

forest_data <- map_df(tests, ~ {
  df <- read_excel("C:/Users/keita/OneDrive - Emory University/Thesis - Anke Huels/Codes/5, 10, 15 year only/Plots Final_LMM_results_all_exposures.xlsx", sheet = .x)
  df$Test <- .x
  df
})

# Clean column names assuming standard format
colnames(forest_data) <- tolower(gsub(" ", "_", colnames(forest_data)))

# Derive exposure variable name (inferred from column index if needed)
forest_data$exposure <- forest_data$...1

# Create plot
forest_plot <- forest_data %>%
  ggplot(aes(x = estimate, y = reorder(exposure, estimate), color = test)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = `2.5_%`, xmax = `97.5_%`), height = 0.2) +
  facet_wrap(~ test, scales = "free_y") +
  labs(
    title = "Association Between PM2.5 Exposure and Cognitive Scores",
    x = "Estimate (95% CI)",
    y = "Exposure Type"
  ) +
  theme_minimal(base_size = 12)

ggplotly(forest_plot, tooltip = c("test", "exposure", "estimate", "2.5_%", "97.5_%"))
```

### Dataset Information {.tabset}

#### Data Description

- **Source**: Emory Healthy Brain Study and SEDAC and RLine Air Pollution Datasets
- **Sample Size**: ~3500 participants
- **Population**: Participants in the Emory Healthy Brain Study in Atlata, GA
- **Time Frame**: 2000–2020

#### Impact Statement

This dashboard illustrates the significant link between air pollution and cognitive decline in older adults. Insights from these findings can support public health interventions and environmental policies.

#### GitHub and Source Code

- [GitHub Repository](https://github.com/its-kei/Data-555-PM2.5-Dementia-Dashboard)

```{r}
# Optionally display the data as a searchable table
DT::datatable(forest_data)
```